---
title: "factor_analysis"
output: pdf_document
---

Here you can learn more about the Data: https://www.worldvaluessurvey.org/WVSDocumentationWV6.jsp

The aim of this work is to highlight specific  interpretation of the democracy in Germany
A theoretical base of this research is written by Inglehart (ect)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(foreign)
library(psych)
library(GGally)
library(GPArotation)
library(nFactors)
library(lavaan)
library(semTools)
library(semPlot)
```

First, read the data in exel formal:

```{r}
#install.packages('readxl')
library(readxl)
dataHW <- read_xlsx(file.choose())
```
PART ONE: pre-processing

Now we choose only whoose columns, which will be used in analysis

```{r}
dataHW <- subset(dataHW, select = c("V131: Democracy: Governments tax the rich and subsidize the poor.", "V132: Democracy: Religious authorities interpret the laws.", "V133: Democracy: People choose their leaders in free elections.", "V134: Democracy: People receive state aid for unemployment.", "V135: Democracy: The army takes over when government is incompetent.", "V136: Democracy: Civil rights protect people’s liberty against oppression.", "V137: Democracy: The state makes people's incomes equal",  "V138: Democracy: People obey their rulers" ,"V139: Democracy: Women have the same rights as men."))
data <- na.omit(dataHW)
summary(data)
```


```{r}
nrow(data)
colnames(data) <- c("v131", "v132", "v133", "v134", "v135", "v136", "v137", "v138", "v139")
describe(data)
```
Leavivg out outliers and measurement errors:
```{r}
all(0 < data)
all(data < 11)
```

```{r}
data <- data[!(data$v131 < 1 | data$v132 < 1 | data$v133 < 1 | data$v134 < 1 | data$v135 < 1 | data$v136 < 1 | data$v137 < 1 | data$v138 < 1 | data$v139 < 1), ]
```

PART TWO: exploratory analysis
Correlation matrix and Kaiser-Meyer statistics
```{r}
corr <- cor(data)
corrplot(corr)
```
Now Im using  Kaiser-Meyer statistics to find out if the data is suttable for factor analysis (tests hypothesis that partial correlations between variables are small - acceptable values are more than 0.6) - all variables are suitable

```{r}
library(psych)
KMO(data)
```

Determining the optimal number of interpretable factors by using Very Simple Structure:
(compared to normal dictribution - ordinate 1)

```{r}
VSS.scree(corr)
```


```{r}
efa1 <- fa(r = corr, nfactors = 2, fm = "pa", rotate = "varimax")
efa1
print(efa1$loadings, cutoff=0.3)

```
```{r}
efa2 <- fa(r = corr, nfactors = 2, fm = "pa", rotate = "oblimin")
efa2

```


```{r}
cfa000 <- 'hard_power=~ v132 + v135 + v138
welfare_state =~ v131 + v134 + v137 
civil_rights =~ v133 + v136 + v139'
model000 <- cfa(cfa000, data)
summary(model000, fit.measures = TRUE, standardized = TRUE)

```

```{r}
cfa1 <- 'radical =~ v132 + v135 + v138
central =~ v131 + v133 + v134 + v136 + v137 + v139'

```

```{r}
library(lavaan)
model1 <- cfa(cfa1, data)
summary(model1, fit.measures = TRUE, standardized = TRUE)

```


```{r}
mepc <- modindices(model1, sort. = TRUE)
mepc
```

```{r}
cfa2 <- 'radical =~ v132 + v135 + v138
central =~ v131 + v133 + v134 + v136 + v137 + v139
v131~~v137'

model2 <- cfa(cfa2, data)
summary(model2, fit.measures = TRUE, standardized = TRUE)

```


```{r}
anova(model1, model2)

```

```{r}
mepc2 <- modindices(model2, sort. = TRUE)
mepc2

```

```{r}
cfa3 <- 'radical =~ v132 + v135 + v138
central =~ v131 + v133 + v134 + v136 + v137 + v139
v131~~v137
v131~~v134'


model3 <- cfa(cfa3, data)
summary(model3, fit.measures = TRUE, standardized = TRUE)

```

```{r}
anova(model2, model3)
```


```{r}
mepc3 <- modindices(model3, sort. = TRUE)
mepc3
```


```{r}
cfa4 <- 'radical =~ v132 + v135 + v138
central =~ v131 + v133 + v134 + v136 + v137 + v139
v131~~v137
v131~~v134
v134~~v137
v133~~v139
v133~~v137
v136~~v137
v137~~v139'

model4 <- cfa(cfa4, data)
summary(model4, fit.measures = TRUE, standardized = TRUE)

```


```{r}
data_2 <- read_xlsx(file.choose())
data_2<- subset(data_2, select = c("V240: Sex", "V131: Democracy: Governments tax the rich and subsidize the poor.", "V132: Democracy: Religious authorities interpret the laws.", "V133: Democracy: People choose their leaders in free elections.", "V134: Democracy: People receive state aid for unemployment.", "V135: Democracy: The army takes over when government is incompetent.", "V136: Democracy: Civil rights protect people’s liberty against oppression.", "V137: Democracy: The state makes people's incomes equal",  "V138: Democracy: People obey their rulers" ,"V139: Democracy: Women have the same rights as men."))
data_2 <- na.omit(data_2)
colnames(data_2) <- c("sex", "v131", "v132", "v133", "v134", "v135", "v136", "v137", "v138", "v139")
summary(data_2)

```

```{r}
configural <- cfa(cfa4, data_2, group="sex")
summary(configural, fit.measures = TRUE)

```


```{r}
metric <- cfa(cfa4, data_2, group="sex",group.equal="loadings")
summary(metric, fit.measures=TRUE)

```

```{r}
CM <- data.frame(cbind(fitMeasures(configural, c("cfi", "rmsea", "srmr")),
                       fitMeasures(metric, c("cfi","rmsea", "srmr"))))
colnames(CM) <- c("Configural", "Metric")
CM$Diff <- CM[,2] - CM[,1]
round(CM, digits = 4)

```


```{r}
anova(configural, metric)
```


```{r}
scalar<- cfa(cfa4, data_2, group="sex", group.equal=c("loadings", "intercepts"))
summary(scalar, fit.measures=TRUE)


anova(metric, scalar)

```

```{r}
MS <- data.frame(cbind(fitMeasures(metric, c("cfi", "rmsea", "srmr")),
                       fitMeasures(scalar, c("cfi","rmsea", "srmr"))))
colnames(MS) <- c("Metric", "Scalar")
MS$Diff <- MS[,2] - MS[,1]
round(MS, digits = 3)

```



Citation:

Inglehart, R., C. Haerpfer, A. Moreno, C. Welzel, K. Kizilova, J. Diez-Medrano, M. Lagos, P. Norris, E. Ponarin & B. Puranen et al. (eds.). 2014. World Values Survey: Round Six - Country-Pooled Datafile Version: https://www.worldvaluessurvey.org/WVSDocumentationWV6.jsp. Madrid: JD Systems Institute.